Visually pleasing knot projections

### 2020-11-04 inst/jma_knot.tex
Journal of Mathematics in the Arts submission kickoff
scale the diagrams so they fit nicely on the page
minor rewriting for JMA

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
knots are said to be equivalent if one can be continuously deformed in
to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Related work in the same language is given by Comoglio
and Rinaldi~\cite{comoglio2011,comoglio2012}; the visual depiction of
knot projections in the context of Celtic artwork is given by
Browne~\cite{browne2005,browne2006}.

@@

The software discussed here~\cite{hankin2016} allows one to specify a
knot in terms of its Bezier control points within inkscape, import the
object into R~\cite{rcore2016}, and then to use numerical optimization
techniques to improve the visual appearance of the knot.

@@

\begin{figure}[htbp]
  \begin{center}
\begin{verbatim}
> k76_rough <- reader(system.file("7_6_first_draft.svg",package="knotR"))
> knotplot2(k76_rough, seg=TRUE)
\end{verbatim}
\begin{centering}
\includegraphics{knot-7_6_rough}
\caption{The \emph{path} of (unoptimized) \label{7_6_rough}
  knot~$7_6$, showing Bezier handles as thin straight lines and
  circles.  The coloured circles have a radius  proportional to the
  curvature (that is, the reciprocal of the radius of curvature) along
  the strand; note large curvature at loop on left}
  \end{centering}
  \end{center}
\end{figure}

\begin{figure}[!tbp]
\begin{verbatim}
knotplot2(k7_6,text=TRUE,lwd=1,circ=0,rainbow=TRUE)
\end{verbatim}
 \centering
\includegraphics{knot-004}
\caption{Knot~$7_6$ with strands numbered \label{k76_strands} to guide
creation of over and under information} \end{figure}

@@

\begin{figure}[!tbp]
\begin{verbatim}
knotplot2(k7_6)
\end{verbatim}
  \centering
\includegraphics{knot-k76_knotplot2}
\caption{Knot $7_6$, post-optimization\label{7_6}}
\end{figure}

@@

\begin{figure}[!tbp]
\begin{verbatim}
knotplot(k7_6)
\end{verbatim}
  \centering
\includegraphics{knot-009}
\caption{Knot $7_6$, post-optimization with breaks indicating
  \label{76_overunder}  underpassing strands}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
\includegraphics[width=13cm]{knots_to_8crossings}
\caption{All \label{all8} prime knots with 8 or fewer crossings,
notation following Rolfsen}
  \end{center}
\end{figure}

>>>>

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
knots are said to be equivalent if one can be continuously deformed in
to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.  In the art world, knot theory
has applications to the study of dance~\cite{khorami2020}, sculpture,
\cite{bosch2010, widmark2020}, and many other branches of mathematical
art~\cite{hart2008}.

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Knot diagrams figure prominently in Celtic
artwork~\cite{cromwell2008} and are discussed from a mathematical
perspective by Browne~\cite{browne2005,browne2006}.

@@

The software discussed here~\cite{hankin2020} allows one to specify a
knot in terms of its Bezier control points within inkscape, import the
object into R~\cite{rcore2016}, and then to use numerical optimization
techniques to improve the visual appearance of the knot.

@@

\begin{figure}[htbp]
  \begin{center}
\begin{verbatim}
> k76_rough <- reader(system.file("7_6_first_draft.svg",package="knotR"))
> knotplot2(k76_rough, seg=TRUE)
\end{verbatim}
\begin{centering}
\includegraphics[scale = 0.9]{knot-7_6_rough}
\caption{The \emph{path} of (unoptimized) \label{7_6_rough}
  knot~$7_6$, showing Bezier handles as thin straight lines and
  circles.  The coloured circles have a radius  proportional to the
  curvature (that is, the reciprocal of the radius of curvature) along
  the strand; note large curvature at loop on left}
  \end{centering}
  \end{center}
\end{figure}

\begin{figure}[!tbp]
\begin{verbatim}
knotplot2(k7_6,text=TRUE,lwd=1,circ=0,rainbow=TRUE)
\end{verbatim}
 \centering
\includegraphics[scale = 0.9]{knot-004}
\caption{Knot~$7_6$ with strands numbered \label{k76_strands} to guide
creation of over and under information} \end{figure}

@@

\begin{figure}[!tbp]
\begin{verbatim}
knotplot2(k7_6)
\end{verbatim}
  \centering
\includegraphics[scale = 0.9]{knot-k76_knotplot2}
\caption{Knot $7_6$, post-optimization\label{7_6}}
\end{figure}

@@

\begin{figure}[!tbp]
\begin{verbatim}
knotplot(k7_6)
\end{verbatim}
  \centering
\includegraphics[scale = 0.9]{knot-009}
\caption{Knot $7_6$, post-optimization with breaks indicating
  \label{76_overunder}  underpassing strands}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
\includegraphics[width=13cm]{knots_to_8crossings}  % created by inst/knotter.R
\caption{All \label{all8} prime knots with 8 or fewer crossings,
notation following Rolfsen}
  \end{center}
\end{figure}
